---
title: "Rmd Exercise Template"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## The question you are answering

You use the `##` to create a header. Use the header to indicate the question you are answering. After the header, in this space here, annotate your answer, telling us what you did.

```{r chunk one}
# Here is where your R code goes. ``` opens the code chunk, followed by {r}. Then ``` closes the chunk. When you 'knit' the file, the codes will evaluate.

##Read the data
dat <- read.csv(here::here("data/gapminder5.csv"), stringsAsFactors=FALSE)
```


## If Statement

Use an if() statement to print a suitable message reporting whether there are any records from 2002 in the gapminder dataset. Now do the same for 2012.

```{r chunk two}
check_for_yearly_data <- function(data, year) 
{
  if (any(data$year == year))
  {
    print(paste("There is data from ", year))
  }
  else
  {
    print(paste("There is NO data from ", year))
  }
}

check_for_yearly_data(dat,2002)
check_for_yearly_data(dat,2012)
```

## Loop and If Statements

Write a script that finds the mean life expectancy by country for countries whose population is below the mean for the dataset

Write a script that loops through the `gapminder` data by continent and prints out whether the mean life expectancy is smaller than 50, between 50 and 70, or greater than 70.

```{r chunk three}
mean_pop = mean(dat$pop,na.rm = T)
dat_sub = subset(dat, dat$pop < mean_pop)
countries <- unique(dat$country)

for (i in countries) 
  {

  mean_le_small_countries <- mean(dat$lifeExp[dat$country == i],na.rm = T)
  print(paste0("Mean life expentancy in ",i,"is: " ,round(mean_le_small_countries,digits =2)))
  }


#Write a script that loops through the gapminder data by continent and prints out whether the mean life expectancy is smaller than 50, between 50 and 70, or greater than 70.
conts <- unique(dat$continent)

for (i in conts) 
{
  mean_le_cont <- mean(dat$lifeExp[dat$continent == i], 
                       na.rm = T)
  
  if (mean_le_cont < 50){
    print(paste0("Mean life expentancy in ",i," is less than 50"))
  }
  
  if (mean_le_cont > 50 & mean_le_cont<70){
    print(paste0("Mean life expentancy in ",i," is greater than 50 but less than 70"))
  }
  else{
    print(paste0("Mean life expentancy in ",i," is greater than 70"))
  }
}

```
## Writing Functions

Create a function that given a data frame will print the name of each column and the class of data it contains.
```{r chunk four}

get_names_classes <- function(data){
  col_names = names(data)
  
  for (i in col_names)
    {
    print(paste0(i," : ",class(i)))
    
    }
}

get_names_classes(dat)
```
Create a function that given a vector will print the mean and the standard deviation of a **vector**, it will optionally also print the median. 
```{r chunk 5}
get_simple_vect_stats<-function(vect,median_flag){
  mean_val = mean(vect, na.rm = T)
  std_val = sd(vect,na.rm=T)
  if (median_flag==T){
    med_val = median(vect,na.rm = T)
    cat("Mean:", mean_val,
        "\nStandard Dev:", std_val,
        "\nMedian:", med_val)
  }
  else{
    cat("Mean:", mean_val,
        "\nStandard Dev:", std_val)
  }
}

le <- dat$lifeExp
get_simple_vect_stats(le,TRUE)
  

```
## ANALYSIS RESULTS
```{r plot chunk}
#Colored by Year
library(ggplot2)
qplot(log(dat$gdpPercap),log(dat$lifeExp),colour = dat$year)
#Colored by Continent
qplot(log(dat$gdpPercap),log(dat$lifeExp),colour = dat$continent)
```

