---
title: "Rmd Exercise Day 2"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T , eval = T,warning = FALSE, message = FALSE,
                      cache = FALSE, tidy = FALSE)
```

## Load the data

Read both California energy datasets. Make sure the `datetime` variable is in an appropriate data type (i.e. not character). 

```{r chunk one}
library(data.table)
library(lubridate)
library(here)
generation <- fread(here::here("data/ca_energy_generation.csv"))
imports <- fread(here::here("data/ca_energy_imports.csv"))
generation$datetime = as_datetime(generation$datetime)
imports$datetime = as_datetime(imports$datetime)

```

## Merge and reshape the data

Merge the two datasets and then melt the resulting dataframe/datatable to make it tidy.

```{r chunk two}
merged_energy <- merge(generation, imports, by = "datetime")
long_merged_energy = melt(merged_energy, id.vars = "datetime",variable.name="source",value.name = "usage")
```

## Creating new variables

Create a series of new variables: 

1. `day`, which is the year-month-day, without the hour. The `lubridate` function `as_date` will do this.
2. `log_output`, which is the natural log of the output.
3. **Challenge**: `per_output`, which is the percent of daily output represented by each observation. You will need to use `group_by` and to create a new variable with the total output for the day. (Make sure to use `ungroup()` after this!)

Bonus: If you are using `dplyr`, try to do this all in one pipe!

```{r chunk three}
long_merged_energy[,day := as_date(datetime)]
long_merged_energy[,log_output := log(usage)]
long_merged_energy[,per_output := usage/sum(usage),by = day]
```

## Summarizing and analyzing data

1. Which source has the greatest mean output by hour? (Hint: Use the `dplyr` verb `arrange(desc(variable))` to order the data frame so that the largest value of `variable` is first. Don't use `desc` and it arranges in ascending order. The `data.table` function is `setorder`.) Which has the least?
2. Which source has the greatest mean output by day? Which has the least? (Do not include zero values.)
3. Which sources has the greatest variance in usage over the course of a dataset? Which has the least? (Do not include zero values.)

```{r chunk four}
#### 1. Which source has the greatest mean output by hour? Which has the least?
new_dt<-long_merged_energy[,.(mean_output_hr = mean(usage,na.rm = T)),by = source]
new_dt[mean_output_hr == max(mean_output_hr) | mean_output_hr == min(mean_output_hr)]
setorder(new_dt,cols=mean_output_hr)
new_dt

#### 2. Which source has the greatest mean output by day? Which has the least? (Do not include zero values.)
new_dt_two<-long_merged_energy[,.(mean_usage_day=mean(usage, na.rm = T)),by = list(source,day)]
setorder(new_dt_two,cols=-mean_usage_day)
head(new_dt_two, n = 10L)

#### 3. Which sources has the greatest variance in usage over the course of a dataset? Which has the least? (Do not include zero values.)
new_dt_three<-long_merged_energy[,.(stdev_usage=sd(usage, na.rm = T)),by = source]
setorder(new_dt_three,cols=-stdev_usage)
new_dt_three
```



## Analyzing renewable versus non-renewable energy sources
```{r chunk five}
regroup <- fread(here::here("data/ca_energy_regroup.csv"))
setnames(regroup,"type","source")
groups_dt <-long_merged_energy[regroup,on="source"]

new_dt_four<-groups_dt[,.(mean_usage=sd(usage, na.rm = T)),by = group]
setorder(new_dt_four,cols=-mean_usage)
new_dt_four
```